---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rctSurv

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

> Efficient Estimators for Survival Analysis in RCTs Without Proportional Hazards

# Installation

The development version can be installed from [GitHub](https://github.com) with:

```{r eval = FALSE}
devtools::install_github("nt-williams/rctSurv")
```

# References

Díaz, I., E. Colantuoni, D. F. Hanley, and M. Rosenblum (2019). Improved precision in
the analysis of randomized trials with survival outcomes, without assuming proportional
hazards. Lifetime Data Analysis 25 (3), 439–468.
