---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# survrct

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![R build status](https://github.com/nt-williams/rctSurv/workflows/R-CMD-check/badge.svg)](https://github.com/nt-williams/rctSurv/actions)
<!-- badges: end -->

> Efficient Estimators for Survival Analysis in RCTs Without Proportional Hazards

Nick Williams and Iván Díaz

---

# Installation

The development version can be installed from [GitHub](https://github.com) with:

```{r eval = FALSE}
devtools::install_github("nt-williams/survrct")
```

# Scope

**survrct** implements efficient estimators for the restricted mean survival time (RMST) and survival probability in randomized controlled trials (RCT) without the proportional hazards assumption as introduced in Diaz et al. (2019). Prognostic baseline variables should be incorporated to obtain equal or better asymptotic precision compared to unadjusted estimators. Under random censoring, the primary estimator (TMLE) is doubly robust--it is consistent if either the outcome or censoring model is correctly specified. This is an advantage over Cox proportional hazards which relies on assumptions of only the outcome model. Estimators are implemented using a formula interface based on that of the [**survival**](https://CRAN.R-project.org/package=survival) package for familiar users.

# Example

To allow for estimation of multiple estimands without having to re-estimate nuisance parameters we first create a Survival metadata object using the `survrct()` function. We specify the model paramaterization using a typical `R` formula with `Surv()` (based on the [**survival**](https://CRAN.R-project.org/package=survival) package) specifying the left-hand side of the formula. 

We also specify the target variable of interest, an optional time coarsener, and the estimator.

```{r eval=FALSE}
library(survrct)

data(colon)
surv <- survrct(Surv(time, status) ~ trt + age + sex + obstruct + perfor + adhere + surg,
                target = "trt", data = colon, coarsen = 30, estimator = "tmle")
#> Surv(time, status) ~ trt + age + sex + obstruct + perfor + adhere + 
#>     surg
#> 
#> ● Estimate RMST with `rmst()`
#> ● Estimate survival probability with `survprob()`
#> ● Inspect nuisance parameter models with `get_fits()`
#> 
#>          Estimator: tmle
#>    Target variable: trt
#>   Status Indicator: status
#>     Adjustment set: age, sex, obstruct, perfor, adhere, and surg
#> Max coarsened time: 111
```

Using the metadata from the previous step we can now estimate the restricted mean survival time and survival probability for a single or multiple time horizons. If multiple times are evaluated, two confidence bands are returned: 95% point-wise intervals as well as 95% uniform confidence bands based on the multiplier-bootstrap from Kennedy (2019).

```{r eval=FALSE}
rmst(surv, 40)
#> RMST Estimator: tmle
#>   Time horizon: 40
#> 
#> Arm-specific RMST:
#> Treatment Arm
#>       Estimate: 31.95
#>     Std. error: 0.73
#>         95% CI: (30.52, 33.38)
#> Control Arm
#>       Estimate: 27.5
#>     Std. error: 0.84
#>         95% CI: (25.86, 29.14)
#> 
#> Treatment Effect:
#> Additive effect
#>       Estimate: 4.45
#>     Std. error: 1.11
#>         95% CI: (2.27, 6.63)
survprob(surv, 40)
#> Survival Probability Estimator: tmle
#>   Time horizon: 40
#> 
#> Arm-specific Survival Probability:
#> Treatment Arm
#>       Estimate: 0.67
#>     Std. error: 0.03
#>         95% CI: (0.62, 0.73)
#> Control Arm
#>       Estimate: 0.53
#>     Std. error: 0.03
#>         95% CI: (0.48, 0.59)
#> 
#> Treatment Effect:
#> Additive effect
#>       Estimate: 0.14
#>     Std. error: 0.04
#>         95% CI: (0.06, 0.22)
```

# References

Díaz, I., E. Colantuoni, D. F. Hanley, and M. Rosenblum (2019). Improved precision in
the analysis of randomized trials with survival outcomes, without assuming proportional
hazards. Lifetime Data Analysis 25 (3), 439–468.

Edward H. Kennedy (2019) Nonparametric Causal Effects Based on Incremental Propensity Score Interventions, Journal of the American Statistical Association, 114:526, 645-656, DOI: 10.1080/01621459.2017.1422737
